Introducing statistics with `intRo`

### 2015-11-23 jcgs-paper-2016/paper.Rmd
jcgs template

---
title: Introducing statistics with \texttt{intRo}
author: Eric Hare & Andee Kaplan
output:
  pdf_document:
    includes:
      in_header: preamble.tex
    number_sections: yes
  html_document: default
bibliography: refs.bib
---

\begin{abstract}
\texttt{intRo} is a modern web-based application for performing basic data analysis and statistical routines. Leveraging the power of \texttt{R} and \texttt{Shiny}, \texttt{intRo} implements common statistical functions in a powerful and extensible modular structure, while remaining simple enough for the novice statistician. This simplicity lends itself to a natural presentation in an introductory statistics course as a substitute for other commonly used statistical software packages, such as \texttt{Excel} and \texttt{JMP}. \texttt{intRo} is currently deployed at the URL \url{http://www.intro-stats.com}. Within this paper, we introduce the application and explore the design decisions underlying \texttt{intRo}, as well as highlight some challenges and advantages of reactive programming.
\end{abstract}

@@

The widespread adoption of \texttt{R} as a tool for statistical analysis has undoubtedly been an important development for the scientific community. However, using \texttt{R} in most cases still requires a basic knowledge of programming concepts which may be unfamiliar to the introductory statistics student. This issue may explain why introductory courses often utilize software applications such as \texttt{JMP}, even though instructors would tend to use a tool like \texttt{R} in their own work. Still, signficant compromises must be made when using software like \texttt{JMP}, including dealing with software licenses, unsupported desktop platforms, and non-ideal user interface decisions.

@@

\texttt{intRo} is a web-based tool to accompany an introductory statistics class. It is meant to assist in the learning of statistics rather than a stand-alone deliverer of statistics education, with the intention of being used in conjunction with a guided class. An accompanying R package, titled \texttt{intRo} and available on GitHub, assists in the downloading, running, and deploying of \texttt{intRo} instances.

@@

# Using intRo
\texttt{intRo} can be downloaded, ran, and deployed on ShinyApps.io through the use of the R package \texttt{intRo}. Currently, the package is only available and GitHub, and can be installed using the \texttt{devtools} package as follows:

```{r, eval=FALSE}
devtools::install_github("gammarama/intRo")
```

After installing the \texttt{intRo} package, the first function one should call is `download_intRo`. `download_intRo` takes as an argument a directory in which to store the application. By default, it selects the working directory of the R session. This function clones the application branch of the \texttt{intRo} repository on GitHub, and hence will pull the latest version of the code whenever it is ran.

Running `download_intRo` will produce an \texttt{intRo} folder in the specified folder. It can then be ran as any \texttt{Shiny} application, using \texttt{Shiny}'s `runApp` command. However, we have provided a wrapper function `run_intRo` which adds some additional customization options to the execution process. `run_intRo` takes as argument the path to the folder containing the \texttt{intRo} application. It also takes several more optional arguments:

* `enabled_modules`: A character vector containing the modules to enable
* `theme`: A string representing a \texttt{shinythemes} theme to use
* `...`: Additional arguments passed to \texttt{Shiny}'s `runApp` function

The package provides help documentation which explains in further detail the format that these arguments would take, but as an example, suppose I wanted to download \texttt{intRo} to my working directory, execute an \texttt{intRo} session with only the data sources, data transform, and numerical summaries modules enabled, and apply the cerulean theme. The series of calls to do so would be as follows:

```{r, eval=FALSE}
download_intRo()
run_intRo(enabled_modules = c("data/transform", "summaries/numerical"), theme = "cerulean")
```

Note that the data sources module is required, and hence must be included in all intRo sessions and need not be specified in the enabled_modules argument.

If the intent is to use a specific instance of \texttt{intRo} where many users will access it at the same time, such as in an introductory statistics class, it may be preferable to deploy a custom instance of \texttt{intRo} to a publicly accessible URL. The package provides a function `deploy_intRo` which is a wrapper for the `deployApp` function contained in the \texttt{shinyapps} package. Once the shinyapps package is installed and configured, `deploy_intRo` will upload \texttt{intRo} as an application on the user's account. The function takes the same arguments as `run_intRo`, so it can be deployed with a custom selection of modules, and a customized theme. It also takes an additional argument `google_analytics`, which allows the specification of a Google Analytics tracking ID. It also takes `...` as additional arguments to be passed into the `deployApp` routine. For example, if we wished to deploy the instance of \texttt{intRo} we ran previously, we would call it like so:

```{r, eval=FALSE}
deploy_intRo(enabled_modules = c("data/transform", "summaries/numerical"), theme = "cerulean")
```

Once the process finished, the app will become available at \url{http://<user>.shinyapps.io/intRo}, where `<user>` is the username of the ShinyApps.io account configured.

>>>>

---
title: "Introducing statistics with `intRo`"
date: "November 23, 2015"
author:
  - name: Eric Hare
    affiliation: Iowa State University
    department: Department of Statistics
  - name: Andee Kaplan
    affiliation: Iowa State University
    department: Department of Statistics
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: preamble.tex
    number_sections: yes
    template: jcgs_article.tex
abstract: |
  `intRo` is a modern web-based application for performing basic data analysis and statistical routines. Leveraging the power of `R` and `Shiny`, `intRo` implements common statistical functions in a powerful and extensible modular structure, while remaining simple enough for the novice statistician. This simplicity lends itself to a natural presentation in an introductory statistics course as a substitute for other commonly used statistical software packages, such as `Excel` and `JMP`. `intRo` is currently deployed at the URL \url{http://www.intro-stats.com}. Within this paper, we introduce the application and explore the design decisions underlying `intRo`, as well as highlight some challenges and advantages of reactive programming.
keywords: blah, blah blah
bibliography: refs.bib
blind: false
---

@@

TODO: Settle on licensing for intRo

The widespread adoption of \texttt{R} as a tool for statistical analysis has undoubtedly beenan important development for the scientific community. However, using \texttt{R} in most cases still requires a basic knowledge of programming concepts which may be unfamiliar to the introductory statistics student. This issue may explain why introductory courses often utilize software applications such as \texttt{JMP}, even though instructors would tend to use a tool like \texttt{R} in their own work. Still, signficant compromises must be made when using software like \texttt{JMP}, including dealing with software licenses, unsupported desktop platforms, and non-ideal user interface decisions.

@@

TODO: Talk about the package (briefly)

\texttt{intRo} is a web-based tool to accompany an introductory statistics class. It is meant to assist in the learning of statistics rather than a stand-alone deliverer of statistics education, with the intention of being used in conjunction with a guided class.

@@

# Using intRo
